---
title: "Logistic Regression"
author: "khalida Khaldi"
date: "1/8/2022"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Logistic Regression Project
###### In this project we will be working with the UCI adult dataset. We will be attempting to predict if people in the data set belong in a certain class by salary, either making <=50k or >50k per year.





```{r}
rm(list = ls()) 

cat("\014")  # ctrl+L

```




## Get the Data

Read in the adult_sal.csv file and set it to a data frame called adult

```{r}
adult <- read.csv('adult_sal.csv')

```

Check the head of adult

```{r}
head(adult)
```


We  notice the index has been repeated. We Drop this column.



```{r}
adult$X <- NULL
head(adult)
```

Check the head,str, and summary of the data now.

```{r}
head(adult)
```



```{r}
str(adult)
```

```{r}
summary(adult)
```


## Data Cleaning

We Notice that we have a lot of columns that are cateogrical factors, however a lot of these columns have too many factors than may be necessary. In this data cleaning section we'll try to clean these columns up by reducing the number of factors.
 
#### type_employer column

To check out the frequency of the type_employer column; we Use table().

```{r}
table(adult$type_employer)

```

How many Null values are there for type_employer? What are the two smallest groups?
We see that  there are 1836 null values, the smallest groups are: Never-worked = 7,  without-pay = 14.



Combine these two smallest groups into a single group called "Unemployed". 



```{r}

func_unpld <- function(work){
    work <- as.character(work)
    if (work=='Never-worked' | work=='Without-pay'){
        return('Unemployed')
    }else{
        return(work)
    }
}

 adult$type_employer <- sapply(adult$type_employer, func_unpld)  



```

Check the  adult$type_employer


```{r}
table(adult$type_employer)

```

#### Marital Column

Use table() to look at the marital column


```{r}
table(adult$marital)
```
Reduce this to three groups:

Married
Not-Married
Never-Married



```{r}
mari_func <- function(statue){
    if (statue=='Divorced' | statue=='Separated' | statue=='Widowed'){
        return('Not-Maried')
      
    }else if(statue=='Never-married'){
        return(statue)
      
    }else{
         return('Married')
    }
}

adult$marital <- sapply(adult$marital, mari_func) 
```


Let us  check the adult$marital



```{r}
table(adult$marital)
```



#### Country Column
Check the country column using table()





```{r message=FALSE, warning=FALSE}
table(adult$country)
```
Group these countries together  by continents. 


```{r message=FALSE, warning=FALSE}
levels(adult$country)
```

```{r message=FALSE, warning=FALSE}
Asia <- c('China','Hong','India','Iran','Cambodia','Japan', 'Laos' ,
          'Philippines' ,'Vietnam' ,'Taiwan', 'Thailand')

North.America <- c('Canada','United-States','Puerto-Rico' )

Europe <- c('England' ,'France', 'Germany' ,'Greece','Holand-Netherlands','Hungary',
            'Ireland','Italy','Poland','Portugal','Scotland','Yugoslavia')

Latin.America <- c('Columbia','Cuba','Dominican-Republic','Ecuador',
                             'El-Salvador','Guatemala','Haiti','Honduras',
                             'Mexico','Nicaragua','Outlying-US(Guam-USVI-etc)','Peru',
                            'Jamaica','Trinadad&Tobago')
Other <- c('unknown')
```


```{r message=FALSE, warning=FALSE}
fun_country <- function(contnt){
    if (contnt %in% Asia){
        return('Asia')
    }else if (contnt %in% North.America){
        return('North.America')
    }else if (contnt %in% Europe){
        return('Europe')
    }else if (contnt %in% Latin.America){
        return('Latin.America')
    }else{
        return('unknown')      
    }
}

```



```{r message=FALSE, warning=FALSE}
adult$country <- sapply(adult$country,fun_country)
```

check adult$country


```{r message=FALSE, warning=FALSE}
table(adult$country)
```

Check the str() of adult again. Make sure any of the columns we changed have factor levels with factor()

```{r message=FALSE, warning=FALSE}
str(adult)
```

convert some columns into factor using factor()

```{r message=FALSE, warning=FALSE}
adult$country <- sapply(adult$country,factor)    # or using factor(), factor(adult$country)
adult$type_employer <- sapply(adult$type_employer,factor)
adult$marital<- sapply(adult$marital,factor)
adult$education<- sapply(adult$education,factor)
adult$occupation <- sapply(adult$occupation ,factor)
adult$relationship <- sapply(adult$relationship ,factor)
adult$race <- sapply(adult$race ,factor)
adult$sex <- sapply(adult$sex ,factor)
adult$income <- sapply(adult$income ,factor)
```

Check adult 
 
 
```{r message=FALSE, warning=FALSE}
str(adult)
```
 
## Missing Data
Install and load the Amelia package.

```{r message=FALSE, warning=FALSE}
library(Amelia)
```

Convert any cell with a '?' or a ' ?' value to a NA value. 

```{r message=FALSE, warning=FALSE}
adult[adult == '?'] <- NA
```

Using table() on a column with NA values. we  see 0 for.


```{r message=FALSE, warning=FALSE}
table(adult$type_employer)
```

```{r message=FALSE, warning=FALSE}
adult$type_employer <- sapply(adult$type_employer,factor)
adult$country <- sapply(adult$country,factor)
adult$marital <- sapply(adult$marital,factor)
adult$occupation <- sapply(adult$occupation,factor)
```


```{r message=FALSE, warning=FALSE}
missmap(adult)
```
Use na.omit() to omit NA data from the adult data frame. 

```{r message=FALSE, warning=FALSE}
adult <- na.omit(adult)
```

Use missmap() to check that all the NA values were in fact dropped.

```{r message=FALSE, warning=FALSE}
 missmap(adult)
```

## Explotory Data Analysis.

 Exploring Data  using visualization.


Use ggplot2 to create a histogram of ages, colored by income.


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```


```{r message=FALSE, warning=FALSE}
ggplot(adult,aes(age)) + geom_histogram(aes(fill=income),color='blue',binwidth=2) + theme_bw()
```
Plot a histogram of hours worked per week



```{r message=FALSE, warning=FALSE}
ggplot(adult,aes(hr_per_week)) + geom_histogram() + theme_bw()
```

Rename the country column to region column to better reflect the factor levels.



```{r message=FALSE, warning=FALSE}
names(adult)[names(adult)=="country"] <- "region"
```

Create a barplot of region with the fill color defined by income class. Optional: Figure out how rotate the x axis text for readability.

```{r message=FALSE, warning=FALSE}
ggplot(adult,aes(region)) + geom_bar(aes(fill=income),color='blue')+theme_bw()+
theme(axis.text.x = element_text( hjust = 2))
```

Take a quick look at the head() of adult to make sure we have a good overview 


```{r message=FALSE, warning=FALSE}
head(adult)
```



Train Test Split
Split the data into a train and test set using the caTools library.


```{r message=FALSE, warning=FALSE}

library(caTools)

set.seed(42) 

sample <- sample.split(adult$income, SplitRatio = 0.80) 
training_feature = subset(adult, sample == TRUE)
test = subset(adult, sample == FALSE)
```

Using  all the features to train a glm() model on the training data set, pass the argument family=binomial(logit) into the glm function.



```{r message=FALSE, warning=FALSE}
model = glm(income ~ ., family = binomial(logit), data = training_feature)
```
Check the model summary


```{r}
##summary(model)
```
Using new.model <- step(your.model.name) to use the step() function to create a new model.


```{r message=FALSE, warning=FALSE}
new.step.model <- step(model)
```

Create a confusion matrix using the predict function with type='response' as an argument inside of that function.




```{r message=FALSE, warning=FALSE}
test$predicted.income = predict(model, newdata=test, type="response")

table(test$income,test$predicted.income > 0.7)
```

