---
title: "Lab_R_5"
author: "khalida"
date: "11/5/2021"
output: html_document
---

Tree Methods Project






```{r}
# Clear environment
rm(list = ls()) 

# Clear packages
# pacman::p_unload(rgl)

# Clear plots
# dev.off()  # But only if there IS a plot

# Clear console
cat("\014")  # ctrl+L
```







#Call the ISLR library , Data for an Introduction to Statistical Learning with Applications in R,
```{r}
install.packages('ISLR')
library(ISLR)
```


#check the head of College 


```{r}
head(College)
```

# reassign College to a dataframe called df

```{r}
df <- College
```


EDA
Let's explore the data!

# Create a scatterplot of Grad.Rate versus Room.Board, colored by the Private column.



```{r}
library(ggplot2)
library(ggthemes)
```

#Create a scatterplot of Grad.Rate versus Room.Board, colored by the Private column.



```{r}
ggplot(data = df, aes(y = Grad.Rate, x = Room.Board, colour = Private)) +
  geom_point()
#color = 'red', size = 1, alpha = 0.6
#ggplot(df,aes(x,y,colour=group))+geom_point()
```
#Create a histogram of full time undergrad students, color by Private.


```{r}
ggplot(data = df, aes(F.Undergrad, fill = Private)) + geom_histogram(color='black')
```
#Create a histogram of Grad.Rate colored by Private. You should see something odd here.


```{r}
ggplot(data = df, aes(Grad.Rate, fill = Private)) + geom_histogram(color='black')

```


#What college had a Graduation Rate of above 100% ?

```{r}
library(dplyr)
```



```{r}

filter(df,Grad.Rate > 100)
```
 
#Change that college's grad rate to 100%

# look for the position of the Grad.Rate =118
```{r}
which(df$Grad.Rate == 118)
```


```{r}
df[92,18] <- 100
```


```{r}
df[92,18] 
```
#Split your data into training and testing sets 70/30. Use the caTools library to do this.

```{r}


library(caTools)

set.seed(123) #

split = sample.split(df$Private, SplitRatio = 0.70)
training_set = subset(df, split == TRUE)
test_set = subset(df, split == FALSE)



```



#Use the rpart library to build a decision tree to predict whether or not a school is Private. Remember to only build your tree off the training 

```{r}


library(rpart)


DTModel <- rpart(formula = Private~ ., method='class',data = training_set)


```

#Use predict() to predict the Private label on the test data.

```{r}

#DTPred <- predict(DTModel, newdata = test_set[-2])
DTPred = predict(DTModel,  test_set)
```

#Check the Head of the predicted values. You should notice that you actually have two columns with the probabilities.



```{r}
head(DTPred)
```

```{r}
df_pred <- as.data.frame(DTPred)
head(df_pred,100)
```


#Turn these two columns into one column to match the original Yes/No Label for a Private column.

```{r}
#df_pred$No
#df_pred$Yes
```





```{r}
df_pred$Private <- NULL
df_pred[df_pred$No > df_pred$Yes,"Private"] <- "No"
df_pred[df_pred$No < df_pred$Yes,"Private"] <- "Yes"
df_pred$Private
```






```{r}
head(df_pred)
```
#Now use table() to create a confusion matrix of your tree model.

```{r}
DTcm = table(df_pred$Private,test_set$Private)
DTcm
```


```{r}
Accuracy_DT = sum(diag(DTcm))/sum(DTcm)
Accuracy_DT



```




#Use the rpart.plot library and the prp() function to plot out your tree model.
```{r}
library(rpart.plot)
prp(DTModel)
```
Call the randomForest package library


```{r}
library(randomForest)
```


#Now use randomForest() to build out a model to predict Private class. Add importance=TRUE as a parameter in the model. (Use help(randomForest) to find out what this does.


```{r}
rf <- randomForest(Private ~ . , data = training_set,importance = TRUE)
```

#What was your model's confusion matrix on its own training set? Use What was your model's confusion matrix on its own training set? Use model$confusion..

```{r}
rf$confusion
```

#Grab the feature importance with model$importance. Refer to the reading for more info on what Gini[1] means.[2]


```{r}
rf$importance
```

