---
title: "TreeModel"
author: "khalidakhaldi"
date: "2022-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tree Method Code:

 For this project I will be exploring the use of tree methods to classify schools as Private or Public based on their features. Let's start by getting the data which is included in the ISLR library, the College data frame.
 
#### A data frame with 777 observations on the following 18 variables:
1. Private A factor with levels No and Yes indicating private or public university
2. Apps Number of applications received
3. Accept Number of applications accepted
4. Enroll Number of new students enrolled
5. Top10perc Pct. new students from top 10% of H.S. class
6. Top25perc Pct. new students from top 25% of H.S. class
7. F.Undergrad Number of fulltime undergraduates
8. P.Undergrad Number of parttime undergraduates
9. Outstate Out-of-state tuition
10.Room.Board Room and board costs
11. Books Estimated book costs
12. Personal Estimated personal spending
13. PhD Pct. of faculty with Ph.D.'s
14. Terminal Pct. of faculty with terminal degree
15. S.F.Ratio Student/faculty ratio
16. perc.alumni Pct. alumni who donate
17. Expend Instructional expenditure per student
18. Grad.Rate Graduation rate








```{r}
# Clear environment
rm(list = ls()) 

# Clear console
cat("\014")  # ctrl+L
```


Call the ISLR library , Data for an Introduction to Statistical Learning with Applications in R
```{r}
#install.packages('ISLR')
library(ISLR)
```

Check the head of College: 

```{r}
head(College)
```


Reassign College to a dataframe called df


```{r}
df <- College
```


##  Explotory Data Anlysis

Let's explore the data!
Creating a scatterplot of Grad.Rate versus Room.Board, colored by the Private column.
```{r message=FALSE, warning=FALSE}
#install.packages("rlang")
```



```{r message=FALSE, warning=FALSE}
library(ggplot2)
#
library(ggthemes)
```


Creating a scatterplot of Grad.Rate versus Room.Board, colored by the Private column.



```{r}
ggplot(data = df, aes(y = Grad.Rate, x = Room.Board, colour = Private)) +
  geom_point()
```


Create a histogram of full time undergrad students, color by Private.


```{r message=FALSE, warning=FALSE}
ggplot(data = df, aes(F.Undergrad, fill = Private)) + geom_histogram(color='black')
```


Creating a histogram of Grad.Rate colored by Private.


```{r message=FALSE, warning=FALSE}
ggplot(data = df, aes(Grad.Rate, fill = Private)) + geom_histogram(color='black')

```

 What college had a Graduation Rate of above 100% ?

```{r message=FALSE, warning=FALSE}
library(dplyr)
```



```{r message=FALSE, warning=FALSE}

filter(df,Grad.Rate > 100)
```
 
looking for the position of the Grad.Rate =118
```{r}
which(df$Grad.Rate == 118)
```


Splitting the  data into training and testing sets 70/30. By  Using the caTools library to do this.

```{r message=FALSE, warning=FALSE}
library(caTools)
set.seed(123) #
split = sample.split(df$Private, SplitRatio = 0.70)
training_set = subset(df, split == TRUE)
test_set = subset(df, split == FALSE)
```


Using the rpart library to build a decision tree to predict whether or not a school is Private.  

```{r}
library(rpart)
DTModel <- rpart(formula = Private~ ., method='class',data = training_set)
```

Using predict() to predict the Private label on the test data.

```{r message=FALSE, warning=FALSE}
DTPred = predict(DTModel,  test_set)
```

Checking the Head of the predicted values. We notice that we actually have two columns with the probabilities.

```{r}
head(DTPred)
```

```{r message=FALSE, warning=FALSE}
df_pred <- as.data.frame(DTPred)
#head(df_pred,100)
```

Turning  the two columns (df_pred(Yes and No)into one column to match the original Yes/No Label for a Private column.

```{r}
df_pred$Private <- NULL
df_pred[df_pred$No > df_pred$Yes,"Private"] <- "No"
df_pred[df_pred$No < df_pred$Yes,"Private"] <- "Yes"
df_pred$Private
```


```{r}
head(df_pred)
```
using table() to create a confusion matrix of your tree model.

```{r}
DTcm = table(df_pred$Private,test_set$Private)
DTcm
```


```{r}
Accuracy_DT = sum(diag(DTcm))/sum(DTcm)
Accuracy_DT



```




Using the rpart.plot library and the prp() function to plot out the  tree model.
```{r}
library(rpart.plot)
prp(DTModel)
```
Call the randomForest package library


```{r message=FALSE, warning=FALSE}
library(randomForest)
```


 Using  randomForest() to build out a model to predict Private class. Adding importance=TRUE as a parameter in the model. 


```{r}
rf <- randomForest(Private ~ . , data = training_set,importance = TRUE)
```

The  model's confusion matrix:

```{r}
rf$confusion
```

Grab the feature importance with model$importance.


```{r}
rf$importance
```

